---
title: "format_check_cell_suppression"
author: "Claire Boulange"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---
```{r}
# Load any necessary packages
library(dplyr)
library(tidyr)
library(readr)
```
Open a connection to the output file
Set the working directory - folder with data cleaned ready for QA

Define a list of acceptable values for the first column
Define a function to check if column names are in snake case format
Define a regex pattern to check age_group format
Define regex patterns for age_group, year_range, and calendar_year

Get a list of CSV files in the working directory
Print the number of files in the folder and their names

initialize an empty list to store data frame
read each CSV file into a data frame and store it in the df_list
append the data frame to the list using the file name as the key

```{r setup, include=FALSE}
# Set the working directory
setwd("C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/ABS BIRTHS///")

first_col_check <- list("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia")

is_snake_case <- function(x) {
  all(grepl("^[a-z]+(_[a-z]+)*$", x))
}

age_group_regex <- "^\\d{1,2}-\\d{1,2}$|^\\d{1,2}-\\d{2}$|^\\d{2}-\\d{2}$"
year_range_regex <- "^\\d{4}-\\d{4}$"
calendar_year_regex <- "^\\d{4}$"


csv_files <- list.files(pattern=".*\\.csv$")

df_list <- list() 

for (file in csv_files) {
  if (endsWith(file, ".csv")) {  # only read in CSV files
    df <- read.csv(file)
    df_list[[file]] <- df
  }
}
```

check for errors in file names

```{r}
cat("Number of files in folder:", length(csv_files), "\n") 

for (file in csv_files) {
  if (!grepl("\\d{3,}", file)) {
    cat("WARNING! File(s) without indicator code:", file, "\n")
  }
}
cat(paste("File with indicator code:",csv_files, collapse = "\n"), "\n")
cat("Finished checking CSV files.", "\n")
```
Loop through each data frame in the list and print the head
```{r}
for (df_name in names(df_list)) {
  cat(paste("Head of", df_name, ":\n"))
  print(head(df_list[[df_name]]))
}
```
print overview of all CSV files and their variables:
```{r}
csv_info <- data.frame(csv_file = names(df_list), variables = sapply(df_list, function(x) paste(names(x), collapse = ", ")))
print(csv_info)
```

This code chunk iterates over each data frame in df_list, retrieves its variable names, class types, range for numeric variables, and unique values for character variables.

It then creates a data frame var_info with these details and prints it, providing a data dictionary for each data frame in the list. You might want to use it to fill the dictionary:

https://connectqutedu.sharepoint.com/:x:/r/teams/FOS_PRO_ANCHDA/Shared%20Documents/General/Metadata/data_dictionary.xlsx?d=w8708a1fa697f42899fd11956f7bc1ce6&csf=1&web=1&e=W3Iqp4

```{r}
# Iterate through each data frame and create a data dictionary
for (df_name in names(df_list)) {
  cat(paste("Data Dictionary for", df_name, ":\n"))
  
  # Get the data frame
  df <- df_list[[df_name]]
  
  # Create a data frame with variable name, class, range, unique values, and count of missing values
  var_info <- data.frame(
    variable = names(df),
    class = sapply(df, class),
    range = sapply(df, function(x) if (is.numeric(x)) paste(range(x, na.rm = TRUE), collapse = " - ") else ""),
    unique_values = sapply(df, function(x) if (is.character(x)) paste(unique(x), collapse = ", ") else ""),
    n_missing_values = sapply(df, function(x) sum(is.na(x)))
  )
  
  # Print the data dictionary
  print(var_info)
}
```

Iterate through each data frame in df_list and perform checks:

1. check if age_group, sex, calendar year or year_range columns exist
2. check if age_group values are in the correct format
3. check if year_range values are in the correct format
4. check if calendar_year values are in the correct format
5. check if geography column is one of the acceptable values
6. print unique values of age_group, calendar_year, year_range and sex

```{r}
for (df_name in names(df_list)) {
  df <- df_list[[df_name]]
  col_names <- names(df)
  first_col <- colnames(df)[1]
  

  if (!("age_group" %in% col_names)) {
    cat("Error: age_group column not found in", df_name, "\n")
  }
  if (!("sex" %in% col_names)) {
    cat("Error: sex column not found in", df_name, "\n")
  }
  if (!("calendar_year" %in% col_names) && !("year_range" %in% col_names)) {
    cat("Error: either calendar_year or year_range column must be present in", df_name, "\n")
  } else if (("calendar_year" %in% col_names) && ("year_range" %in% col_names)) {
    cat("Error: both calendar_year and year_range columns cannot be present in", df_name, "\n")
  } else if ("calendar_year" %in% col_names && !all(grepl("\\d{4}", df$calendar_year))) {
    cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
  } else if ("year_range" %in% col_names && !all(grepl("\\d{4}-\\d{4}", df$year_range))) {
    cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
  }
  

  if ("age_group" %in% col_names) {
    age_group_values <- df$age_group
    if (!all(grepl(age_group_regex, age_group_values))) {
      cat("Error: age_group values in", df_name, "are not in the correct format (expected format: \\d-\\d, \\d-\\d{2}, or \\d{2}-\\d{2})\n")
      cat("Invalid values:\n")
      invalid_age_group_values <- age_group_values[!grepl(age_group_regex, age_group_values)]
      cat(paste(unique(invalid_age_group_values), collapse=", "), "\n")
    }
  }

  if ("year_range" %in% col_names) {
    year_range_values <- df$year_range
    if (!all(grepl(year_range_regex, year_range_values))) {
      cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
      cat("Invalid values:\n")
      invalid_year_range_values <- year_range_values[!grepl(year_range_regex, year_range_values)]
      cat(paste(unique(invalid_year_range_values), collapse=", "), "\n")
    }
  }

  if ("calendar_year" %in% col_names) {
    calendar_year_values <- df$calendar_year
    if (!all(grepl(calendar_year_regex, calendar_year_values))) {
      cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
      cat("Invalid values:\n")
      invalid_calendar_year_values <- calendar_year_values[!grepl(calendar_year_regex, calendar_year_values)]
      cat(paste(unique(invalid_calendar_year_values), collapse=", "), "\n")
    }
  }
  
  if (!is_snake_case(col_names[-1])) {
    cat("Error: column names in", df_name, "are not in snake case format (lowercase words separated by underscores)\n")
  }

  if (!(first_col %in% first_col_check)) {
    cat("Error: geography column is not one of the acceptable values (", paste(first_col_check, collapse=", "), ") in", df_name, "\n")
  }
  
  if (first_col != "Australia" && any(df[[1]] == 0)) {
  cat("Error: Values coded as 0 (Australia) found in a dataset that is not national:", df_name, "\n")
  }
}
```
CHECK OUTPUTS NOW!!!

Did you detect any formatting errors? 
-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
YES >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> GO BACK TO YOUR CODE AND MAKE CORRECTIONS
-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
NO  >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> PROCEED WITH CELL SUPPRESSION
-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

Define the input directory path
Define the output directory path to save cleaned dataset WITH CELL SUPPRESSED
```{r}

input_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/ABS BIRTHS//"
output_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/ABS BIRTHS//cell_suppressed/"
```

Create the output directory if it doesn't already exist
Get a list of all CSV files in the input directory

```{r}
dir.create(output_dir, showWarnings = FALSE)
csv_files <- list.files(input_dir, pattern = ".csv$", full.names = TRUE)
```
Loop through each CSV file 
Check if there is an "uncertainty" column in the data frame
```{r}
for (file in csv_files) {
  
 
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  
  if ("uncertainty" %in% colnames(df)) {
    message(paste0("Note: The file ", basename(file), " contains an 'uncertainty' column. Make sure to modify the code to apply cell suppression where uncertainty is high."))}
  else {
  # Print message indicating that there is no need to apply cell suppression to "uncertainty" column
  cat("You don't have to worry about cell suppression on 'uncertainty' in", basename(file), "\n")
}
}

```
detect columns that are numeric and where you might need to apply cell suppression
```{r}
# Define the exclusion list
exclude_list <- c("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia", "sex", "calendar_year", "age_group", "year_range", "irsd_quintile")

# Loop through each CSV file and check for columns that are numeric and not in the exclusion list
for (file in csv_files) {
  
  # Read in the CSV file
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  # Get the names of columns that are numeric and not in the exclusion list
  num_cols <- names(df)[sapply(df, is.numeric) & !names(df) %in% exclude_list]
  
  # If there are any such columns, print a message for each file and column
  if (length(num_cols) > 0) {
    for (col in num_cols) {
      message(paste0("For file ", basename(file), ", check values in column '", col, "' for cell suppression"))
    }
  } else {
    # Print message indicating that there are no columns to check
    cat("You don't have to worry about cell suppression in any numeric columns in", basename(file), "\n")
  }
  
}
```
Loop through each CSV file and apply cell suppression
```{r}
for (file in csv_files) {
  
  # Read in the CSV file
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  # Check if the n_births column exists in the data frame
  if ("n_births" %in% colnames(df)) {

    df[df[,"n_births"] %in% 1:4 & !is.na(df[,"n_births"]), -c(1:4)] <- 9999999          # modify this if you copy this code or need to change the cell suppression rules
  } else {
    # Print a message to indicate that the column was not found
    cat("Skipping file", file, "because it does not contain the n_births column.\n")
  }
  
  # Write the modified data frame to a new CSV file in the output directory
  write.csv(df, file.path(output_dir, basename(file)), row.names = FALSE)
  
}
```

