---
title: "format check and cell suppression on NDSH data"
author: "Claire Boulange"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---
```{r}
# Load any necessary packages
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
library(kableExtra)
```
Part 1:
- open a connection to the output file
- Set the working directory - folder with data cleaned ready for QA

- define a list of acceptable values for the first column
- define a function to check if column names are in snake case format
- define a regex pattern to check age_group format
- define regex patterns for age_group, year_range, and calendar_year
- get a list of CSV files in the working directory + subfolders
- print the number of files in the folder and their names

- initialize an empty list to store data frame
- read each CSV file into a data frame and store it in the df_list
- append the data frame to the list using the file name as the key

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = 'markdown')

# Set the working directory
setwd("C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/NDSHS/") #change this path if you copy this 
csv_files <- list.files(pattern=".*\\.csv$")
first_col_check <- list("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia")

is_snake_case <- function(x) {
  all(grepl("^[a-z]+(_[a-z]+)*$", x))
}

age_group_regex <- "^\\d{1,2}-\\d{1,2}$|^\\d{1,2}-\\d{2}$|^\\d{2}-\\d{2}$"
year_range_regex <- "^\\d{4}-\\d{4}$"
calendar_year_regex <- "^\\d{4}$"

df_list <- list() 

for (file in csv_files) {
  if (endsWith(file, ".csv")) {  # only read in CSV files
    df <- read.csv(file)
    df_list[[file]] <- df
  }
}
```

```{r}
cat("Number of files in folder:", length(csv_files), "\n") 

for (file in csv_files) {
  if (!grepl("\\d{3,}", file)) {
    cat("WARNING! File(s) without indicator code:", file, "\n")
  }
}
cat(paste("File with indicator code:",csv_files, collapse = "\n"), "\n")
cat("Finished checking CSV files.", "\n")

for (file in csv_files) {
  if (!grepl("_STE|_SA3|_SA2|_SA4|_national|_Australia", file)) {
    cat("WARNING! File(s) without geography suffix:", file, "\n")
  }
}

cat(paste("File with indicator code:", csv_files, collapse = "\n"), "\n")
cat("Finished checking CSV files.", "\n")

```
Part 2
- loop through each data frame in the list and print the head
```{r, results='asis'}
for (df_name in names(df_list)) {
  cat(paste("Head of", df_name, ":\n"))
  
  # Get the data frame
  df <- df_list[[df_name]]
  
  # Generate the HTML table for the head of the data frame
  html_table <- kable(head(df), format = "html") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
  
  # Print the HTML table
  cat(as.character(html_table))
  cat("\n")
}


```
- print overview of all CSV files and their variables
```{r, results='asis'}


# Create a new data frame with only the variables column
csv_info <- data.frame(variables = sapply(df_list, function(x) paste(names(x), collapse = ", ")))

# Generate the HTML table using kable and kableExtra functions
html_table <- kable(csv_info, format = "html", col.names = NULL) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Print the HTML table
cat(as.character(html_table))


```
- print a data dictionary for each data frame
```{r, results='asis'}
library(knitr)
library(kableExtra)

# Loop for generating data dictionary
for (df_name in names(df_list)) {
  cat(paste("Data Dictionary for", df_name, ":\n"))
  
  # Get the data frame
  df <- df_list[[df_name]]
  
  # Create a data frame with variable name, class, range, unique values, and count of missing values
  var_info <- data.frame(
    variable = names(df),
    class = sapply(df, class),
    range = sapply(df, function(x) if (is.numeric(x)) paste(range(x, na.rm = TRUE), collapse = " - ") else ""),
    unique_values = sapply(df, function(x) if (is.character(x)) paste(unique(x), collapse = ", ") else ""),
    n_missing_values = sapply(df, function(x) sum(is.na(x)))
  )
  
  # Generate the HTML table using kable and kableExtra functions
  html_table <- kable(var_info, format = "html") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)

  # Print the HTML table
  cat(as.character(html_table))
  cat("\n")
}

```
Part 3 - iterate through each data frame in df_list and perform checks:

1. check if age_group, sex, calendar year or year_range columns exist
2. check if age_group values are in the correct format
3. check if year_range values are in the correct format
4. check if calendar_year values are in the correct format
5. check if geography column is one of the acceptable values


```{r}
for (df_name in names(df_list)) {
  df <- df_list[[df_name]]
  col_names <- names(df)
  first_col <- colnames(df)[1]
  

  if (!("age_group" %in% col_names)) {
    cat("Error: age_group column not found in", df_name, "\n")
  }
  if (!("sex" %in% col_names)) {
    cat("Error: sex column not found in", df_name, "\n")
  }
  if (!("calendar_year" %in% col_names) && !("year_range" %in% col_names)) {
    cat("Error: either calendar_year or year_range column must be present in", df_name, "\n")
  } else if (("calendar_year" %in% col_names) && ("year_range" %in% col_names)) {
    cat("Error: both calendar_year and year_range columns cannot be present in", df_name, "\n")
  } else if ("calendar_year" %in% col_names && !all(grepl("\\d{4}", df$calendar_year))) {
    cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
  } else if ("year_range" %in% col_names && !all(grepl("\\d{4}-\\d{4}", df$year_range))) {
    cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
  }
  

  if ("age_group" %in% col_names) {
    age_group_values <- df$age_group
    if (!all(grepl(age_group_regex, age_group_values))) {
      cat("Error: age_group values in", df_name, "are not in the correct format (expected format: \\d-\\d, \\d-\\d{2}, or \\d{2}-\\d{2})\n")
      cat("Invalid values:\n")
      invalid_age_group_values <- age_group_values[!grepl(age_group_regex, age_group_values)]
      cat(paste(unique(invalid_age_group_values), collapse=", "), "\n")
    }
  }

  if ("year_range" %in% col_names) {
    year_range_values <- df$year_range
    if (!all(grepl(year_range_regex, year_range_values))) {
      cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
      cat("Invalid values:\n")
      invalid_year_range_values <- year_range_values[!grepl(year_range_regex, year_range_values)]
      cat(paste(unique(invalid_year_range_values), collapse=", "), "\n")
    }
  }

  if ("calendar_year" %in% col_names) {
    calendar_year_values <- df$calendar_year
    if (!all(grepl(calendar_year_regex, calendar_year_values))) {
      cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
      cat("Invalid values:\n")
      invalid_calendar_year_values <- calendar_year_values[!grepl(calendar_year_regex, calendar_year_values)]
      cat(paste(unique(invalid_calendar_year_values), collapse=", "), "\n")
    }
  }
  
  if (!is_snake_case(col_names[-1])) {
    cat("Error: column names in", df_name, "are not in snake case format (lowercase words separated by underscores)\n")
  }

  if (!(first_col %in% first_col_check)) {
    cat("Error: geography column is not one of the acceptable values (", paste(first_col_check, collapse=", "), ") in", df_name, "\n")
  }
  
  if (!is.na(first_col) && first_col != "Australia" && any(df[[1]] == 0, na.rm = TRUE)) {
  cat("Error: Values coded as 0 (Australia) found in a dataset that is not national:", df_name, "\n")
  }
}
```
                                                                                          STOP HERE AND CHECK!!!

Did you detect any formatting errors?

YES >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> GO BACK TO YOUR CODE AND MAKE CORRECTIONS

NO  >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> PROCEED WITH CELL SUPPRESSION


\\
Part 4 - cell suppression
- define the input directory path
- define the output directory path to save cleaned datasets WITH CELL SUPPRESSED (R will automatically create this folder if it does not exist)
```{r}
input_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/NDSHS/" #change this path
output_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/NDSHS/cell_suppressed/"#change this path
```
run this code to...
- create the output directory if it doesn't already exist
- get a list of all CSV files in the input directory
```{r}
dir.create(output_dir, showWarnings = FALSE)
csv_files <- list.files(input_dir, pattern = ".csv$", full.names = TRUE)
```
Run this code to... 
check if there is an "uncertainty" column in the data frame
```{r}
for (file in csv_files) {
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  if (any(grepl("uncertainty", colnames(df)))) {
    cat("Note: The file ", basename(file), " contains a column with the word 'uncertainty' in its name. Make sure to modify the code to apply cell suppression where necessary.\n")
  } else {
    cat("You don't have to worry about cell suppression on 'uncertainty' in", basename(file), "\n")
  }
}

```
Run this code to...
detect columns that are numeric and where you might need to apply cell suppression
```{r}
cat("Checking for columns requiring cell suppression:\n\n")

# Define the exclusion list
exclude_list <- c("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia", "sex", "calendar_year", "age_group", "year_range", "irsd_quintile")

# Define the pattern to match column names
pattern <- "^n\\_|number|count"

# Loop through each CSV file and check for columns that are numeric and not in the exclusion list and have names matching the pattern
for (file in csv_files) {
  
  # Read in the CSV file
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  # Get the names of columns that are numeric and not in the exclusion list and have names matching the pattern
  num_cols <- names(df)[sapply(df, is.numeric) & !names(df) %in% exclude_list & grepl(pattern, names(df), ignore.case = TRUE)]
  
  # If there are any such columns, print a message for each file and column
  if (length(num_cols) > 0) {
    for (col in num_cols) {
      message(paste0("For file ", basename(file), ", check values in column '", col, "' for cell suppression"))
    }
  } else {
    # Print message indicating that there are no columns to check
    cat("You don't have to worry about cell suppression in any relevant columns in", basename(file), "\n")
  }
  
}

```


```{r}
cat("Checking for columns with values 1, 2, 3, or 4:\n\n")

# Define the pattern to match column names
pattern <- "^n_"

# Define the values to check for
values_to_check <- c(1, 2, 3, 4)

# Variable to keep track of whether values are found
values_found <- FALSE

# Variable to store column names with values in the specified range
columns_with_values <- character()

# Loop through each CSV file and check for columns with names matching the pattern
for (file in csv_files) {

  # Read in the CSV file
  df <- read.csv(file, stringsAsFactors = FALSE)

  # Get the names of columns with names matching the pattern
  matching_cols <- names(df)[grepl(pattern, names(df))]

  # Loop through each matching column and check for values 1, 2, 3, or 4
  for (col in matching_cols) {
    if (any(df[[col]] %in% values_to_check)) {
      message(paste0("For file ", basename(file), ", values 1, 2, 3, or 4 found in column '", col, "'"))
      values_found <- TRUE
      columns_with_values <- c(columns_with_values, col)
    }
  }

}

# If no values in the specified range are found, print a message
if (!values_found) {
  cat("No values 1, 2, 3, or 4 found in columns with 'n_' prefix.\n")
} else {
  cat("Columns with values 1, 2, 3, or 4 in columns with 'n_' prefix:\n")
  cat(columns_with_values, "\n")
}

```


Once you understand what you need to do with cell suppression (which columns represent count values in your series of data set AND if you have uncertainty columns to deal with) customise the code below to apply cell suppression  

```{r}
# List the CSV files in the directory
csv_files <- list.files(input_dir, pattern = "*.csv", full.names = TRUE)

for (file in csv_files) {
  
  # Read the CSV file
  
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  
  
  # Get the column names
  
  column_names <- colnames(df)
  
  
  
  # Loop through each column
  
  for (i in 2:length(column_names)) {
    
    # Check if the column matches the pattern "uncertainty"
    
    if (grepl("_uncertainty", column_names[i])) {
      
      suppressed_column <- i - 1
      
      #looping through each cell
      
      for(j in 1:nrow(df)){
        # Check if the value in the uncertainty column is 2
        if (!is.na(df[[i]][j]) && df[[i]][j] == 2) {
          
          # Suppress the column on the left of the uncertainty column
          
          df[j, suppressed_column] <- 9999999
          
          print(paste("Suppressed column", column_names[suppressed_column]))
          
        }
        
      }
      print(paste("Suppressed column", column_names[suppressed_column]))
      
      
    }
    
  }
  
  
  # Remove columns containing the word "uncertainty"
  
  df <- df[, !grepl("uncertainty", column_names)]
  
  
  
  # Write the modified data frame to a new CSV file in the output directory
  
  write.csv(df, file.path(output_dir, basename(file)), row.names = FALSE)
  
  
  
  print(paste("Modified file", basename(file), "has been written"))
  
}


```

