---
title: "format_check_cell_suppression"
author: "Claire Boulange"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---
```{r}
# Load any necessary packages
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
library(kableExtra)
```
Part 1:
- open a connection to the output file
- Set the working directory - folder with data cleaned ready for QA

- define a list of acceptable values for the first column
- define a function to check if column names are in snake case format
- define a regex pattern to check age_group format
- define regex patterns for age_group, year_range, and calendar_year
- get a list of CSV files in the working directory + subfolders
- print the number of files in the folder and their names

- initialize an empty list to store data frame
- read each CSV file into a data frame and store it in the df_list
- append the data frame to the list using the file name as the key
```{r setup, include=FALSE}
# Set the working directory
setwd("C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/nsshs/") #change this path if you copy this 
csv_files <- list.files(pattern=".*\\.csv$")
first_col_check <- list("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia")

is_snake_case <- function(x) {
  all(grepl("^[a-z]+(_[a-z]+)*$", x))
}

age_group_regex <- "^\\d{1,2}-\\d{1,2}$|^\\d{1,2}-\\d{2}$|^\\d{2}-\\d{2}$"
year_range_regex <- "^\\d{4}-\\d{4}$"
calendar_year_regex <- "^\\d{4}$"

df_list <- list() 

for (file in csv_files) {
  if (endsWith(file, ".csv")) {  # only read in CSV files
    df <- read.csv(file)
    df_list[[file]] <- df
  }
}
```
```{r}
cat("Number of files in folder:", length(csv_files), "\n") 

for (file in csv_files) {
  if (!grepl("\\d{3,}", file)) {
    cat("WARNING! File(s) without indicator code:", file, "\n")
  }
}
cat(paste("File with indicator code:",csv_files, collapse = "\n"), "\n")
cat("Finished checking CSV files.", "\n")

for (file in csv_files) {
  if (!grepl("_STE|_SA3|_SA2|_SA4|_national|_Australia", file)) {
    cat("WARNING! File(s) without geography suffix:", file, "\n")
  }
}

cat(paste("File with indicator code:", csv_files, collapse = "\n"), "\n")
cat("Finished checking CSV files.", "\n")

```
Part 2
loop through each data frame in the list and print the head
```{r, results='asis'}
for (df_name in names(df_list)) {
  cat(paste("Head of", df_name, ":\n"))
  
  # Get the data frame
  df <- df_list[[df_name]]
  
  # Generate the HTML table for the head of the data frame
  html_table <- kable(head(df), format = "html") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
  
  # Print the HTML table
  cat(as.character(html_table))
  cat("\n")
}


```
- print overview of all CSV files and their variables
```{r, results='asis'}


# Create a new data frame with only the variables column
csv_info <- data.frame(variables = sapply(df_list, function(x) paste(names(x), collapse = ", ")))

# Generate the HTML table using kable and kableExtra functions
html_table <- kable(csv_info, format = "html", col.names = NULL) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Print the HTML table
cat(as.character(html_table))


```
- print a data dictionary for each data frame
```{r, results='asis'}
# Loop for generating data dictionary
for (df_name in names(df_list)) {
  cat(paste("Data Dictionary for", df_name, ":\n"))
  
  # Get the data frame
  df <- df_list[[df_name]]
  
  # Create a data frame with variable name, class, range, unique values, count of missing values, and count of rows
  var_info <- data.frame(
    variable = names(df),
    class = sapply(df, class),
    range = sapply(df, function(x) if (is.numeric(x)) paste(range(x, na.rm = TRUE), collapse = " - ") else ""),
    unique_values = sapply(df, function(x) if (is.character(x)) paste(unique(x), collapse = ", ") else ""),
    n_missing_values = sapply(df, function(x) sum(is.na(x))),
    n_rows = nrow(df)
  )
  
  # Generate the HTML table using kable and kableExtra functions
  html_table <- kable(var_info, format = "html") %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
  
  # Print the HTML table
  cat(as.character(html_table))
  cat("\n")
}

```
Part 3 - iterate through each data frame in df_list and perform checks:
1. check if age_group, sex, calendar year or year_range columns exist
2. check if age_group values are in the correct format
3. check if year_range values are in the correct format
4. check if calendar_year values are in the correct format
5. check if geography column is one of the acceptable values
```{r}
for (df_name in names(df_list)) {
  df <- df_list[[df_name]]
  col_names <- names(df)
  first_col <- colnames(df)[1]
  

  if (!("age_group" %in% col_names)) {
    cat("Error: age_group column not found in", df_name, "\n")
  }
  if (!("sex" %in% col_names)) {
    cat("Error: sex column not found in", df_name, "\n")
  }
  if (!("calendar_year" %in% col_names) && !("year_range" %in% col_names)) {
    cat("Error: either calendar_year or year_range column must be present in", df_name, "\n")
  } else if (("calendar_year" %in% col_names) && ("year_range" %in% col_names)) {
    cat("Error: both calendar_year and year_range columns cannot be present in", df_name, "\n")
  } else if ("calendar_year" %in% col_names && !all(grepl("\\d{4}", df$calendar_year))) {
    cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
  } else if ("year_range" %in% col_names && !all(grepl("\\d{4}-\\d{4}", df$year_range))) {
    cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
  }
  

  if ("age_group" %in% col_names) {
    age_group_values <- df$age_group
    if (!all(grepl(age_group_regex, age_group_values))) {
      cat("Error: age_group values in", df_name, "are not in the correct format (expected format: \\d-\\d, \\d-\\d{2}, or \\d{2}-\\d{2})\n")
      cat("Invalid values:\n")
      invalid_age_group_values <- age_group_values[!grepl(age_group_regex, age_group_values)]
      cat(paste(unique(invalid_age_group_values), collapse=", "), "\n")
    }
  }

  if ("year_range" %in% col_names) {
    year_range_values <- df$year_range
    if (!all(grepl(year_range_regex, year_range_values))) {
      cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
      cat("Invalid values:\n")
      invalid_year_range_values <- year_range_values[!grepl(year_range_regex, year_range_values)]
      cat(paste(unique(invalid_year_range_values), collapse=", "), "\n")
    }
  }

  if ("calendar_year" %in% col_names) {
    calendar_year_values <- df$calendar_year
    if (!all(grepl(calendar_year_regex, calendar_year_values))) {
      cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
      cat("Invalid values:\n")
      invalid_calendar_year_values <- calendar_year_values[!grepl(calendar_year_regex, calendar_year_values)]
      cat(paste(unique(invalid_calendar_year_values), collapse=", "), "\n")
    }
  }
  
  if (!is_snake_case(col_names[-1])) {
    cat("Error: column names in", df_name, "are not in snake case format (lowercase words separated by underscores)\n")
  }

  if (!(first_col %in% first_col_check)) {
    cat("Error: geography column is not one of the acceptable values (", paste(first_col_check, collapse=", "), ") in", df_name, "\n")
  }
  
  if (!is.na(first_col) && first_col != "Australia" && any(df[[1]] == 0, na.rm = TRUE)) {
  cat("Error: Values coded as 0 (Australia) found in a dataset that is not national:", df_name, "\n")
  }
}
```
                                                                                          STOP HERE AND CHECK!!!

Did you detect any formatting errors?

YES >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> GO BACK TO YOUR CODE AND MAKE CORRECTIONS

NO  >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> PROCEED WITH CELL SUPPRESSION

Part 4 - cell suppression
- define the input directory path
- define the output directory path to save cleaned datasets WITH CELL SUPPRESSED (R will automatically create this folder if it does not exist)
```{r}

input_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/nsshs//"
output_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/nsshs/cell_suppressed/"
```
Create the output directory if it doesn't already exist
Get a list of all CSV files in the input directory

```{r}

dir.create(output_dir, showWarnings = FALSE)
csv_files <- list.files(input_dir, pattern = ".csv$", full.names = TRUE)
```
Loop through each CSV file 
Check if there is an "uncertainty" column in the data frame

```{r}
for (file in csv_files) {
  
 
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  
  if ("uncertainty" %in% colnames(df)) {
    message(paste0("Note: The file ", basename(file), " contains an 'uncertainty' column. Make sure to modify the code to apply cell suppression where uncertainty is high."))}
  else {
  # Print message indicating that there is no need to apply cell suppression to "uncertainty" column
  cat("You don't have to worry about cell suppression on 'uncertainty' in", basename(file), "\n")
}
}
```
detect columns that are numeric and where you might need to apply cell suppression
```{r}
# Define the exclusion list
exclude_list <- c("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia", "sex", "calendar_year", "age_group", "year_range", "irsd_quintile")

# Loop through each CSV file and check for columns that are numeric and not in the exclusion list
for (file in csv_files) {
  
  # Read in the CSV file
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  # Get the names of columns that are numeric and not in the exclusion list
  num_cols <- names(df)[sapply(df, is.numeric) & !names(df) %in% exclude_list]
  
  # If there are any such columns, print a message for each file and column
  if (length(num_cols) > 0) {
    for (col in num_cols) {
      message(paste0("For file ", basename(file), ", check values in column '", col, "' for cell suppression"))
    }
  } else {
    # Print message indicating that there are no columns to check
    cat("You don't have to worry about cell suppression in any numeric columns in", basename(file), "\n")
  }
  
}
```
Loop through each CSV file and apply cell suppression
Write the modified data frame to a new CSV file in the output directory

```{r}

```

