---
title: "format_check_cell_suppression"
author: "Claire Boulange"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
# Open a connection to the output file
# Set the working directory - folder with data cleaned ready for QA
```{r}
setwd("C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/NHS/")
```

# Define a list of acceptable values for the first column
# Define a function to check if column names are in snake case format
# Define a regex pattern to check age_group format
# Define regex patterns for age_group, year_range, and calendar_year
```{r}
first_col_check <- list("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia")

is_snake_case <- function(x) {
  all(grepl("^[a-z]+(_[a-z]+)*$", x))
}

age_group_regex <- "^\\d{1,2}-\\d{1,2}$|^\\d{1,2}-\\d{2}$|^\\d{2}-\\d{2}$"
year_range_regex <- "^\\d{4}-\\d{4}$"
calendar_year_regex <- "^\\d{4}$"
```

# Get a list of CSV files in the working directory + subfolders
# Print the number of files in the folder and their names
```{r}
csv_files <- list.files(pattern=".*\\.csv$", recursive = TRUE)
cat("Number of files in folder:", length(csv_files), "\n") 
cat(paste(csv_files, collapse = "\n"), "\n")

```

initialize an empty list to store data frame
read each CSV file into a data frame and store it in the df_list
append the data frame to the list using the file name as the key
```{r}
df_list <- list() 


for (file in csv_files) {
  if (endsWith(file, ".csv")) {  # only read in CSV files
    df <- read.csv(file)
    df_list[[file]] <- df
  }
}

```

# Iterate through each data frame in df_list and perform checks:

1. check if age_group, sex, calendar year or year_range columns exist
2. check if age_group values are in the correct format
3. check if year_range values are in the correct format
4. check if calendar_year values are in the correct format
5. check if geography column is one of the acceptable values
6. print unique values of age_group, calendar_year, year_range and sex

```{r}

for (df_name in names(df_list)) {
  df <- df_list[[df_name]]
  col_names <- names(df)
  first_col <- colnames(df)[1]
  

  if (!("age_group" %in% col_names)) {
    cat("Error: age_group column not found in", df_name, "\n")
  }
  if (!("sex" %in% col_names)) {
    cat("Error: sex column not found in", df_name, "\n")
  }
  if (!("calendar_year" %in% col_names) && !("year_range" %in% col_names)) {
    cat("Error: either calendar_year or year_range column must be present in", df_name, "\n")
  } else if (("calendar_year" %in% col_names) && ("year_range" %in% col_names)) {
    cat("Error: both calendar_year and year_range columns cannot be present in", df_name, "\n")
  } else if ("calendar_year" %in% col_names && !all(grepl("\\d{4}", df$calendar_year))) {
    cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
  } else if ("year_range" %in% col_names && !all(grepl("\\d{4}-\\d{4}", df$year_range))) {
    cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
  }
  


  if ("age_group" %in% col_names) {
    age_group_values <- df$age_group
    if (!all(grepl(age_group_regex, age_group_values))) {
      cat("Error: age_group values in", df_name, "are not in the correct format (expected format: \\d-\\d, \\d-\\d{2}, or \\d{2}-\\d{2})\n")
    }
  }
  

  if ("year_range" %in% col_names) {
    year_range_values <- df$year_range
    if (!all(grepl(year_range_regex, year_range_values))) {
      cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
    }
  }

  if ("calendar_year" %in% col_names) {
    calendar_year_values <- df$calendar_year
    if (!all(grepl(calendar_year_regex, calendar_year_values))) {
      cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
    }
  }
  
  if (!is_snake_case(col_names[-1])) {
    cat("Error: column names in", df_name, "are not in snake case format (lowercase words separated by underscores)\n")
  }

  if (!(first_col %in% first_col_check)) {
    cat("Error: geography column is not one of the acceptable values (", paste(first_col_check, collapse=", "), ") in", df_name, "\n")
  }
  

  cat("Unique values in age_group for", df_name, ":\n")
  cat(unique(df$age_group), "\n\n")
  
  cat("Unique values in calendar_year for", df_name, ":\n")
  cat(unique(df$calendar_year), "\n\n")
  
  cat("Unique values in year_range for", df_name, ":\n")
  cat(unique(df$year_range), "\n\n")
  
  cat("Unique values in sex for", df_name, ":\n")
  cat(unique(df$sex), "\n\n")
}

```
#CHECK OUTPUTS NOW - GO BACK TO YOUR CODE IF ANY FORMATTING ERRORS DETECTED

All good? great! now you can move on to the cell suppression 


Define the input directory path
Define the output directory path to save cleaned dataset WITH CELL SUPPRESSED
```{r}

input_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/SHSC/"


output_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/SHSC//cell_suppressed/"
```

Create the output directory if it doesn't already exist
Get a list of all CSV files in the input directory
Loop through each CSV file and apply cell suppression (here we are searching in the 5th column for values < 5)


```{r}

dir.create(output_dir, showWarnings = FALSE)


csv_files <- list.files(input_dir, pattern = ".csv$", full.names = TRUE)


for (file in csv_files) {
  
  # Read in the CSV file
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  # Apply cell suppression
  df[df[,"client_count_SHSC"] %in% 0:4 & !is.na(df[,"client_count_SHSC"]), -c(1:4)] <- 9999999

  
  # Write the modified data frame to a new CSV file in the output directory
  write.csv(df, file.path(output_dir, basename(file)), row.names = FALSE)
  
}
```

