---
title: "format_check_cell_suppression"
author: "Claire Boulange"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
# Open a connection to the output file
# Set the working directory - folder with data cleaned ready for QA
```{r}
setwd("C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/abs_income")
```

# Define a list of acceptable values for the first column
# Define a function to check if column names are in snake case format
# Define a regex pattern to check age_group format
# Define regex patterns for age_group, year_range, and calendar_year
```{r}
first_col_check <- list("STE_CODE16", "SA2_CODE16", "SA3_CODE16", "SA4_CODE16", "LGA_CODE16", "Australia")

is_snake_case <- function(x) {
  all(grepl("^[a-z]+(_[a-z]+)*$", x))
}

age_group_regex <- "^\\d{1,2}-\\d{1,2}$|^\\d{1,2}-\\d{2}$|^\\d{2}-\\d{2}$"
year_range_regex <- "^\\d{4}-\\d{4}$"
calendar_year_regex <- "^\\d{4}$"
```

# Get a list of CSV files in the working directory
# Print the number of files in the folder and their names
```{r}
csv_files <- list.files(pattern=".*\\.csv$")
cat("Number of files in folder:", length(csv_files), "\n") 
cat(paste(csv_files, collapse = "\n"), "\n")

```

initialize an empty list to store data frame
read each CSV file into a data frame and store it in the df_list
append the data frame to the list using the file name as the key
```{r}
df_list <- list() 


for (file in csv_files) {
  if (endsWith(file, ".csv")) {  # only read in CSV files
    df <- read.csv(file)
    df_list[[file]] <- df
  }
}

```

# Iterate through each data frame in df_list and perform checks:

1. check if age_group, sex, calendar year or year_range columns exist
2. check if age_group values are in the correct format
3. check if year_range values are in the correct format
4. check if calendar_year values are in the correct format
5. check if geography column is one of the acceptable values
6. print unique values of age_group, calendar_year, year_range and sex

```{r}

for (df_name in names(df_list)) {
  df <- df_list[[df_name]]
  col_names <- names(df)
  first_col <- colnames(df)[1]
  

  if (!("age_group" %in% col_names)) {
    cat("Error: age_group column not found in", df_name, "\n")
  }
  if (!("sex" %in% col_names)) {
    cat("Error: sex column not found in", df_name, "\n")
  }
  if (!("calendar_year" %in% col_names)) {
    cat("Error: calendar_year column not found in", df_name, "\n")
  }
  if (!("year_range" %in% col_names)) {
    cat("Error: year_range column not found in", df_name, "\n")
  }
  


  if ("age_group" %in% col_names) {
    age_group_values <- df$age_group
    if (!all(grepl(age_group_regex, age_group_values))) {
      cat("Error: age_group values in", df_name, "are not in the correct format (expected format: \\d-\\d, \\d-\\d{2}, or \\d{2}-\\d{2})\n")
    }
  }
  

  if ("year_range" %in% col_names) {
    year_range_values <- df$year_range
    if (!all(grepl(year_range_regex, year_range_values))) {
      cat("Error: year_range values in", df_name, "are not in the correct format (expected format: \\d{4}-\\d{4})\n")
    }
  }

  if ("calendar_year" %in% col_names) {
    calendar_year_values <- df$calendar_year
    if (!all(grepl(calendar_year_regex, calendar_year_values))) {
      cat("Error: calendar_year values in", df_name, "are not in the correct format (expected format: \\d{4})\n")
    }
  }
  
  if (!is_snake_case(col_names[-1])) {
    cat("Error: column names in", df_name, "are not in snake case format (lowercase words separated by underscores)\n")
  }

  if (!(first_col %in% first_col_check)) {
    cat("Error: geography column is not one of the acceptable values (", paste(first_col_check, collapse=", "), ") in", df_name, "\n")
  }
  

  cat("Unique values in age_group for", df_name, ":\n")
  cat(unique(df$age_group), "\n\n")
  
  cat("Unique values in calendar_year for", df_name, ":\n")
  cat(unique(df$calendar_year), "\n\n")
  
  cat("Unique values in year_range for", df_name, ":\n")
  cat(unique(df$year_range), "\n\n")
  
  cat("Unique values in sex for", df_name, ":\n")
  cat(unique(df$sex), "\n\n")
}

```
##CHECK OUTPUTS NOW - GO BACK TO YOUR CODE IF ANY FORMATTING ERRORS DETECTED

```{r}
# Define the input directory path
input_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/abs_income/"

# Define the output directory path
output_dir <- "C:/Users/00095998/OneDrive - The University of Western Australia/acwa_temp/abs_income//cell_suppressed/"
```


```{r}
# Create the output directory if it doesn't already exist
dir.create(output_dir, showWarnings = FALSE)

# Get a list of all CSV files in the input directory
csv_files <- list.files(input_dir, pattern = ".csv$", full.names = TRUE)

# Loop through each CSV file and apply cell suppression
for (file in csv_files) {
  
  # Read in the CSV file
  df <- read.csv(file, stringsAsFactors = FALSE)
  
  # Apply cell suppression
  df[df[, 5] < 5, -c(1:4)] <- 9999999
  
  # Write the modified data frame to a new CSV file in the output directory
  write.csv(df, file.path(output_dir, basename(file)), row.names = FALSE)
  
}
```

