---
title: "Fixing Census Variables"
author: "Owen Forbes"
date: "26/04/2023"
output: html_document
---

# Added at 19/05/23
- update file path to new directory for data that has had temporal correspondence applied, before name fixes
- "Australia" --> "0" for Unemployment

# Cleaning up Census datasets - fixed column names and variables.

Each code chunk loops over the 6 geographies for one census dataset.


```{r setup, include=FALSE}
library(tidyverse)


census_geog_types <- c("LGA", "SA2", "SA3", "SA4", "State", "Australia")
```




Fixing Unemployment - keeping only n_young_people_unemployed_seeking_full_time_work

```{r}

fixing_unemployment <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_unemployment_",geog,".csv"))
  
  names(temp_data)[6] <- "n_young_people_unemployed_seeking_full_time_work" #rename variable
  
  temp_data <- temp_data %>% filter(lfhrp_labour_force_status == "Unemployed, looking for full-time work") #keep only this level
  
  temp_data <- temp_data[,-4] #drop old/un-needed filter column
  
  if(geog == "Australia"){temp_data$Australia <- rep("0", length(temp_data[,1]))}
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_251_unemployed_and_seeking_full_time_work_",geog,".csv"), row.names = F)
}
  
  
for (geog_i in census_geog_types) {
  fixing_unemployment(geog_i)
}
  
```



Fixing Internet connection type - renaming cols  internet_connection_type +  n_private_dwellings 

```{r}



fixing_internet <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_internet_connection_",geog,".csv"))
  
  names(temp_data)[2] <- "internet_connection_type"
  
  names(temp_data)[3] <- "n_private_dwellings"
  
  temp_data <- temp_data[,c(1,2,4,3)] #put calendar year before variable
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_282_internet_connection_at_home_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_internet(geog_i)
}
```



Fixing Tenure type - renaming cols  tenure_type and n_private_dwellings

```{r}

fixing_tenure <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_tenure_type_",geog,".csv"))
  
  names(temp_data)[2] <- "household_tenure_type"
  
  names(temp_data)[4] <- "n_private_dwellings"
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_283_household_tenure_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_tenure(geog_i)
}
```



Fixing One Parent Families - keeping cols n_of_one_parent_families_with_non_dependent_children_only and n_of_one_parent_families_with_children_under_15_years_old

```{r}

fixing_oneparent <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_one_parent_",geog,".csv"))
  
  
  names(temp_data)[4] <- "n_of_one_parent_families_with_children_under_15_years_old"
  
  temp_data <- temp_data %>% filter(fmcf_family_composition == "One parent family with children under 15")
  
  # remove filter col
  
  temp_data <- temp_data[,-2]
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_321_one_parent_families_with_children_under_15_years_old_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_oneparent(geog_i)
}
```



Fixing Year 12 - n_young_people_with_year_12_or_equivalent_highest_year_of_school_completed

```{r}

fixing_yr12 <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_year12_",geog,".csv"))
  
  names(temp_data)[6] <- "n_young_people_with_year_12_or_equivalent_highest_year_of_school_completed"
  
  temp_data <- temp_data[,-4] #remove old filter col
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_462_highest_year_of_school_completed_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_yr12(geog_i)
}
```






Fixing Volunteering - n_young_people_doing_unpaid_voluntary_work

```{r}

fixing_volunteer <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_volunteer_",geog,".csv"))
  
  names(temp_data)[6] <- "n_young_people_doing_unpaid_voluntary_work"
  
  temp_data <- temp_data[,-4] #remove old filter col
  
  temp_data <- temp_data %>% filter(age_group == "15-19" | age_group == "20-24") #remove age groups 0-14 - no volunteering recorded in TB data
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_551_volunteering_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_volunteer(geog_i)
}
```




Fixing LOTE spoken at home

```{r}

fixing_LANP <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_LANP_",geog,".csv"))
  
  names(temp_data)[4] <- "language_group_spoken_at_home"
  
 names(temp_data)[6] <- "n"  
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_632_children_and_young_people_who_speak_a_language_other_than_english_at_home_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_LANP(geog_i)
}
```






Fixing - English spoken not well or not at all

```{r}

fixing_english <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_english_notwell_",geog,".csv"))
  

  temp_data <- temp_data %>% group_by_at(c(1,2,3,5)) %>% mutate(new_vals = sum(n_children_and_young_people_who_speak_english_not_well_or_not_at_all))
  
  #temp_data <- temp_data %>% group_by(.data[[geog]], age_group, sex, calendar_year) %>% mutate(new_vals = sum(n_children_and_young_people_who_speak_english_not_well_or_not_at_all))
  
  temp_data[,6] <- temp_data[,7]
  
  names(temp_data)[6] <- "n_children_and_young_people_who_speak_english_not_well_or_not_at_all"
  
  
  # remove unnecessary columns
  temp_data <- temp_data[,-c(4,7)]
  
  # keep unique rows
  
  temp_data <- distinct(temp_data, .keep_all = T)
  
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_633_children_and_young_people_who_speak_english_not_well_or_not_at_all_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_english(geog_i)
}
```







Fixing country/region of birth - region_children_and_young_people_born + n

```{r}

fixing_country <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_countryofbirth_",geog,".csv"))
  
  names(temp_data)[4] <- "region_children_and_young_people_born"
  
  names(temp_data)[6] <- "n"
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_661_country_of_birth_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_country(geog_i)
}
```




Fixing Religion - religious_affiliation + n

```{r}

fixing_religion <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_religion_",geog,".csv"))
  
  names(temp_data)[4] <- "religious_affiliation"
  
  names(temp_data)[6] <- "n"
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_671_religious_affiliation_of_15_24_year_olds_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_religion(geog_i)
}
```




Fixing core activity need for assistance - n_children_and_young_people_with_core_activity_need_for_assistance

```{r}

fixing_core <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_core_activity_need_assistance_",geog,".csv"))
  
  
  temp_data <- temp_data %>% filter(assnp_core_activity_need_for_assistance=="Has need for assistance with core activities")
  
  names(temp_data)[6] <- "n_children_and_young_people_with_core_activity_need_for_assistance"
  
  temp_data <- temp_data[,-4]
  
  # Remove "Total" entries for state
  if(geog == "State"){temp_data <- filter(temp_data, State != "Total")}

  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_1121_children_with_core_activity_need_for_assistance_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_core(geog_i)
}
```



Fixing same sex households - n_same_sex_couples_with_one_or_more_children
n_same_sex_couples_with_no_children

```{r}

fixing_ssh <- function(geog){
  
  temp_data <- read.csv(paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_INTERIM/Census_Interim_Pre-Temporal-Concordance/after_correspondence_before_name_fixes/census_same_sex_couple_",geog,".csv"))
  
  temp_data <- temp_data %>% filter(cdcf_count_of_dependent_children_in_family!="No children" & cdcf_count_of_dependent_children_in_family!="Not applicable" & cdcf_count_of_dependent_children_in_family!="Couple family with: No dependent children")
  
  temp_data <- temp_data %>% filter(sscf_same_sex_couple_indicator=="Male same sex couple" | sscf_same_sex_couple_indicator=="Female same sex couple" | sscf_same_sex_couple_indicator=="Female same-sex couple" | sscf_same_sex_couple_indicator=="Male same-sex couple")
  
  
  

  temp_data <- temp_data %>% group_by_at(c(1,4)) %>% mutate(new_vals = sum(same_sex_couple_children))
  
  temp_data[,5] <- temp_data[,6]
  
  names(temp_data)[5] <- "n_same_sex_couples_with_one_or_more_dependent_children"
  
  
  # remove unnecessary columns
  temp_data <- temp_data[,-c(2,3,6)]
  
  # keep unique rows
  
  temp_data <- distinct(temp_data, .keep_all = T)
  
  write.csv(temp_data, file = paste0("/Users/Current/OneDrive - Queensland University of Technology/General - ACWA_QUT/Data_Collections_READY_FOR_QA/census/ABS_census_323_same_sex_parents_with_dependent_children_",geog,".csv"), row.names = F)
}
  
  

for (geog_i in census_geog_types) {
  fixing_ssh(geog_i)
}
```
